Mining of Keystroke and Mouse Dynamics to Increase the Engagement of Students with Programming Assignments

### 2018-04-21 paper.tex
Puts parameters in table form

    \end{tabular}
  \end{table}

As part of the data mining process a canonical genetic algorithm was used for feature
selection with the parameters used shown in Table \ref{tab_ga}. These
parameters were set heuristically, and were the same for selecting the
features of all classification algorithms. % Mario check this - JJ
In the end, the subset for the frustration classifier includes only 11 features, and 
the subset for the boredom classifier consists of only 13 features.

% It's not clear what's the fitness function here or what it really does. Maybe we could expand this part of the work, test different evolutionary algorithms and write a paper on feature selection using evolutionary algorithms for EvoStar - JJ

The selected classifiers where implemented using Rapid Miner, with the
following parameters: For J48 decision tree algorithm a confidence threshold
for pruning of 0.25, with a minimum of 2 instances per leaf, and 3 folds for
reduced error pruning. The feed-forward neural network used a back-propagation
learning algorithm with two hidden layer of 20 neurons each, with a learning
rate and momentum of 0.6, and was trained for 200 generations.

Table \ref{tab_performance} shows the performance of each of the four
classifiers together
with the $\kappa$ coefficients. A 10-Fold cross validation was used.
The accuracies and $\kappa$ coefficients obtained are close to what is
usually obtained in fixed-text methods, for example, the results presented in
Epp, Lippold y Mandryk \cite{epp2011identifying}.
In a fixed-text method, subjects are asked to write specific texts.
The results obtained with this method are satisfactory, considering that
learners were writing a program, a task comparable with free-text writing,
where there is no restriction on what has to be written.
In the case of the $\kappa$ coefficients, it is
usual to see values below 0.2 in methods involving free-text. In this case, some
values of $\kappa$ were close to 0.5, is important to consider the values of $\kappa$
in these results because the class distribution is not uniformly distributed. As
it is observed in many works in affective recognition, decision trees normally
produce competitive accuracy. In this case the J-48 classifier obtained an
accuracy of 80.48\%, which is marginally better than the others, and an accepted $\kappa$
statistic for the kind of problem. The artificial neural network (ANN)
gives the highest
accuracy in the classification of frustration.
%
\begin{table}[h!tb]
\centering
\caption{Parameters used in the evolutionary algorithm for feature
  weight optimization.}
\label{tab_ga_w}
    \begin{tabular}{ | l | l | }
      \hline
      Parameter & Value \\
        \hline
      Population size & 10 \\
      Generations & 30 \\
      Selection & Tournament \\
      Tournament size & population size / 4 \\
      Crossover & Uniform \\
      Crossover probability & 0.5 \\
      Mutation rate & $1 \over 39$ \\
      Minimum number of features & 5 \\
        \hline
    \end{tabular}
  \end{table}
% 
\begin{table*}[!t]
\centering
\caption{Parameters used for the different classification algorithms }
\label{tab_params_ml}
\label{tab_perf_flow}
    \begin{tabular}{ | l | l |  }

      Algorithm & Parameters \\
      \hline
    \end{tabular}
  \end{table}

Table \ref{tab_perf_flow} shows the performance for the additional classifiers
evaluated specifically for the Flow/engaged affective state. The goal was to
increase the performance achieved earlier. Again a ten-
fold cross-validation was conducted. In this case, instead of using an
evolutionary algorithm to optimize feature selection, it was used for 
optimizing a weight vector. 
      The feature-weight vector assigned a weight to each one of the
      features, and the parameters that have been used in the
      evolutionary algorithm are shown in Table \ref{tab_ga_w}.
      % The
% parameters used for the evolutionary optimization algorithm are the following:
% Population size = 10, generations = 30, selection = tournament, tournament size = 0.25 (fraction of the population),  initialize probability = 0.5,  mutation
%       probability = 1/number of attributes,  crossover probability =
%       0.5.
      % Please Mario check these parameters and how they have put to
      % the table; I don't know what's the "initialization
      % probability" - JJ

      Parameters
used for the Random Forest algorithm: number of trees = 10, 
criterion = gain ratio, maximal depth = 20, 
confidence = 0.25, minimal gain = 0.1, minmail leaf size = 2.  The
parameters used for are the J48 are the following:  confidence = 0.79, minimal
leaf size=2. For the kNN algorithm, the value k = 4 was used. 
For gradient boosted
trees: number of trees = 20,  maximal depth = 5, minimum rows = 10, 
minimum split 
improvement = 0.0, number of bins = 20, 
learning rate = 0.1, sample rate = 1.0.  Bayes
Kernel: laplace correction = True, estimation mode = greedy, 
minimum bandwidth = 0.1, number of kernels = 10.  Finally for deep learning: activation = Rectifier, hidden layer sizes = [50, 50], epochs = 10.

>>>>

    \end{tabular}
  \end{table}
  %
  \begin{table*}[h!tb]
\centering
\caption{Parameters used in the classification algorithms.}
\label{tab_clas}
\begin{tabular}{ | l | p{10cm} | }
  \hline
  Algorithm & Parameters \\
  \hline
  J48 & Confidence threshold for pruning = 0.25; minimum instances per
        leaf = 2; folds for reduced error pruning = 3 \\
  Neural network & Learning algorithm: Back-propagation; hidden layers
                   = 2, hidden layer size = 20 neurons; learning rate
                   = 0.6; momentum = 0.6; training iterations = 200 \\
  \hline
\end{tabular}
\end{table*}
%
As part of the data mining process a canonical genetic algorithm was used for feature
selection with the parameters used shown in Table \ref{tab_ga}. These
parameters were set heuristically, and were the same for selecting the
features of all classification algorithms. % Mario check this - JJ
In the end, the subset for the frustration classifier includes only 11 features, and 
the subset for the boredom classifier consists of only 13 features. 
%
% It's not clear what's the fitness function here or what it really does. Maybe we could expand this part of the work, test different evolutionary algorithms and write a paper on feature selection using evolutionary algorithms for EvoStar - JJ
%
The selected classifiers where implemented using Rapid Miner, with the
  parameters shown in Table \ref{tab_clas}.

%   For J48 decision tree algorithm a confidence threshold
% for pruning of 0.25, with a minimum of 2 instances per leaf, and 3 folds for
% reduced error pruning. The feed-forward neural network used a back-propagation
% learning algorithm with two hidden layer of 20 neurons each, with a learning
% rate and momentum of 0.6, and was trained for 200 generations.

Table \ref{tab_performance} shows the performance of each of the four
classifiers together
with the $\kappa$ coefficients. A 10-Fold cross validation was used.
The accuracies and $\kappa$ coefficients obtained are close to what is
usually obtained in fixed-text methods, for example, the results presented in
Epp, Lippold y Mandryk \cite{epp2011identifying}.
In a fixed-text method, subjects are asked to write specific texts.
The results obtained with this method are satisfactory, considering that
learners were writing a program, a task comparable with free-text writing,
where there is no restriction on what has to be written.
In the case of the $\kappa$ coefficients, it is
usual to see values below 0.2 in methods involving free-text. In this case, some
values of $\kappa$ were close to 0.5, is important to consider the values of $\kappa$
in these results because the class distribution is not uniformly distributed. As
it is observed in many works in affective recognition, decision trees normally
produce competitive accuracy. In this case the J-48 classifier obtained an
accuracy of 80.48\%, which is marginally better than the others, and an accepted $\kappa$
statistic for the kind of problem. The artificial neural network (ANN)
gives the highest
accuracy in the classification of frustration.
%
\begin{table}[h!tb]
\centering
\caption{Parameters used in the evolutionary algorithm for feature
  weight optimization.}
\label{tab_ga_w}
    \begin{tabular}{ | l | l | }
      \hline
      Parameter & Value \\
        \hline
      Population size & 10 \\
      Generations & 30 \\
      Selection & Tournament \\
      Tournament size & population size / 4 \\
      Crossover & Uniform \\
      Crossover probability & 0.5 \\
      Mutation rate & $1 \over 39$ \\
      Minimum number of features & 5 \\
        \hline
    \end{tabular}
  \end{table}
% 
\begin{table*}[!t]
\centering
\caption{Parameters used for the different classification algorithms
  used for recognizing the Flow/engaged state.}
\label{tab_params_ml}
    \begin{tabular}{ | l | p{10cm} |  }
\hline
      Algorithm & Parameters \\
      \hline 
      Random Forest &  number of trees = 10, 
criterion = gain ratio, maximal depth = 20, 
                      confidence = 0.25, minimal gain = 0.1, minmail leaf size = 2 \\
      J48 &  confidence = 0.79, minimal leaf size=2 \\
      kNN & $k$ = 4 \\
      Gradient boosted trees & number of trees = 20,  maximal depth = 5, minimum rows = 10, 
minimum split 
improvement = 0.0, number of bins = 20, 
                               learning rate = 0.1, sample rate = 1.0 \\
      Bayes Kernel & laplace correction = True, estimation mode = greedy, 
                     minimum bandwidth = 0.1, number of kernels = 10 \\
      Deep learning & activation = Rectifier, hidden layer sizes =
                      [50, 50], epochs = 10 \\
      \hline
    \end{tabular}
  \end{table*}

Table \ref{tab_perf_flow} shows the performance for the additional classifiers
evaluated specifically for the Flow/engaged affective state. The goal was to
increase the performance achieved earlier. Again a ten-
fold cross-validation was conducted. In this case, instead of using an
evolutionary algorithm to optimize feature selection, it was used for 
optimizing a weight vector. 
      The feature-weight vector assigned a weight to each one of the
      features, and the parameters that have been used in the
      evolutionary algorithm are shown in Table \ref{tab_ga_w}.
      % The
% parameters used for the evolutionary optimization algorithm are the following:
% Population size = 10, generations = 30, selection = tournament, tournament size = 0.25 (fraction of the population),  initialize probability = 0.5,  mutation
%       probability = 1/number of attributes,  crossover probability =
%       0.5.
      % Please Mario check these parameters and how they have put to
      % the table; I don't know what's the "initialization
      % probability" - JJ
%
      The parameters used in this case for the classification
      algorithms, and the algorithms used are shown in Table \ref{tab_params_ml}
% used for the Random Forest algorithm: number of trees = 10, 
% criterion = gain ratio, maximal depth = 20, 
% confidence = 0.25, minimal gain = 0.1, minmail leaf size = 2.  The
% parameters used for are the J48 are the following:  confidence = 0.79, minimal
% leaf size=2. For the kNN algorithm, the value k = 4 was used. 
% For gradient boosted
% trees: number of trees = 20,  maximal depth = 5, minimum rows = 10, 
% minimum split 
% improvement = 0.0, number of bins = 20, 
% learning rate = 0.1, sample rate = 1.0.  Bayes
% Kernel: laplace correction = True, estimation mode = greedy, 
% minimum bandwidth = 0.1, number of kernels = 10.  Finally for deep learning: activation = Rectifier, hidden layer sizes = [50, 50], epochs = 10.
